---
title-block-banner: true
title: "Analysis of PDGFR-Î²-low and PDGFR-Î² high+ cells in the ipsilateral hemisphere (striatal-only lesions)"
subtitle: "Data analysis notebook"
date: today
date-format: full
author: 
  - name: "Daniel Manrique-Castano"
    orcid: 0000-0002-1912-1764
    degrees:
      - PhD
    affiliation: 
      - name: Univerisity Laval 
        department: Psychiatry and Neuroscience
        group: Laboratory of neurovascular interactions 
note: "GitHub: https://daniel-manrique.github.io/"

keywords: 
  - PDGFR-Î²
  - Brain injury
  - Bayesian modeling 
   
license: "CC BY"

format:
   pdf: 
    toc: true
    number-sections: true
    colorlinks: true
   html:
    code-fold: true
    embed-resources: true
    toc: true
    toc-depth: 2
    toc-location: left
    theme: spacelab

knitr:
  opts_chunk: 
    warning: false
    message: false

csl: science.csl
bibliography: references.bib
---

# Preview

In this notebook, we analyze the proportion and distribution of PDGFR-Î²_Low and PDGFR-Î²_High+ in brains with only striatal lesions. We performed a separate image pre-processing, cell detection and data analysis for these animals given that the experiments were carried out at a different moment. We performed automatic cell detection and classification using QuPath [@bankhead2017]. We also create point patterns using the coordinates of detecting cells to perform Point Pattern analysis.

**Parent dataset:** PDGFR-Î² and GFAP-stained ischemic hemispheres imaged at 5x (with stitching). Samples are grouped at 14 and 30 days post-ischemia (DPI). The raw images and pre-processing scripts (if applicable) are available at the Zenodo repository (10.5281/zenodo.10553084) under the name `Widefield_5x_Ipsilateral_Gfap-Pdgfrb.zip`. Individual cells were detected and classified into PDGFR-Î²^low^ (Pdgfrb_NonReact) and PDGFR-Î²^high^ (Pdgfrb_React) using QuPath [@bankhead2017].The complete QuPath project, including classifiers and output data as .tsv files is available at https://osf.io/ty4z5.

**Working dataset**: The `Widefield_5x_Ipsilateral_Gfap-Pdgfrb_Inten_Str.csv`data frame containing the number of PDGFR-Î²^low^ (Pdgfrb_NonReact) and PDGFR-Î²^high^ (Pdgfrb_React) cells in the ischemic hemisphere. Here, we analyze the proportion and distribution of these populations.

# Install and load required packages

Install and load all required packages. Please uncomment (delete #) the line code if installation is required. Load the installed libraries each time you start a new R session.

```{r}
#| label: Install_Packages
#| include: true
#| warning: false
#| message: false

#install.packages("devtools")
#library(devtools)

#install.packages("devtools")
#library(devtools)

#install.packages(c("bayesplot", "bayestestR", "brms","broom.mixed", "dplyr", "easystats", "distributional", "ggplot","gtsummary", "modelbased", "modelr", "modelsummary", "patchwork", "poorman", "tidybayes", "tidyverse"))

library(bayesplot)
library(bayestestR)
library(brms)
library(broom.mixed)
library(dplyr)
library(easystats)
library(distributional)
library(ggplot2)
library(gtsummary)
library(modelbased)
library(modelr)
library(modelsummary)
library(patchwork)
library(poorman)
library(tidybayes)
library(tidyverse)
```

# Visual themes

We create a visual theme to use in our plots (ggplots).

```{r}
#| label: Plot_Theme
#| include: true
#| warning: false
#| message: false
  
Plot_theme <- theme_classic() +
  theme(
      plot.title = element_text(size=18, hjust = 0.5, face="bold"),
      plot.subtitle = element_text(size = 10, color = "black"),
      plot.caption = element_text(size = 12, color = "black"),
      axis.line = element_line(colour = "black", size = 1.5, linetype = "solid"),
      axis.ticks.length=unit(7,"pt"),
     
      axis.title.x = element_text(colour = "black", size = 16),
      axis.text.x = element_text(colour = "black", size = 16, angle = 0, hjust = 0.5),
      axis.ticks.x = element_line(colour = "black", size = 1),
      
      axis.title.y = element_text(colour = "black", size = 16),
      axis.text.y = element_text(colour = "black", size = 16),
      axis.ticks.y = element_line(colour = "black", size = 1),
      
      legend.position="right",
      legend.direction="vertical",
      legend.title = element_text(colour="black", face="bold", size=12),
      legend.text = element_text(colour="black", size=10),
      
      plot.margin = margin(t = 10,  # Top margin
                             r = 2,  # Right margin
                             b = 10,  # Bottom margin
                             l = 10) # Left margin
      ) 
```

# Exploratory data visualization

```{r}
#| label: Pdgfrb_Striatum_Table
#| include: true
#| warning: false
#| message: false
#| tbl-cap: "Data set"

# We load the dataset in case is not present in the R environment
Pdgfrb_Striatum <- read.csv(file = "ProData/Wide5x_Ipsi_Pdgfr-Gfap_LowHigh_Str.csv", header = TRUE)

knitr::kable(Pdgfrb_Striatum [1:10, ])
```

For this analysis, we focus on `DPI` (Days post-ischemia), `Pdgfr_NonReact`, `Pdgfr_React` variables to analyze these cells proportions in the ischemic brain. Next, we visualize the raw data to guide the statistical modeling. We plot the response variables as a scatter plot fitting a linear model, given that this data set comprises just two time points.

```{r}
#| label: Pdgfrb_LowHigh_Str_Exploratory
#| include: true
#| warning: false
#| message: false
#| results: false
#| fig-cap: Exploratory data visualization for PDGFR-Î²^low^ and PDGFR-Î²^high^ (striatal lesions)
#| fig-width: 9
#| fig-height: 4


set.seed(8807)

# PDGFR-Î²^low^
###################

Striatum_Pdgfr_Low_Sctr <- 
  ggplot(
    data  = Pdgfrb_Striatum, 
    aes(x = DPI, 
        y = Pdgfr_NonReact)) +
geom_smooth(
  method = "lm", 
  se     = TRUE,
  color  = "black") +
geom_jitter(
  width = 0.5, 
  shape = 1, 
  size  = 1.5, 
  color = "black") +
  scale_y_continuous(name= expression("Number of PDGFR-Î²"^low)) +
  scale_x_continuous(name="Days post-ischemia (DPI) ",
                     breaks=c(14,30)) +
  Plot_theme
  

# PDGFR-Î²^low^
#######################

Striatum_Pdgfr_High_Sctr <- 
  ggplot(
    data  = Pdgfrb_Striatum, 
    aes(x = DPI, 
        y = Pdgfr_React)) +
geom_smooth(
  method = "lm", 
  se     = TRUE,
  color  = "black") +
geom_jitter(
  width = 0.5, 
  shape = 1, 
  size  = 1.5, 
  color = "black") +
  scale_y_continuous(name= expression("Number of PDGFR-Î²"^high)) +
  scale_x_continuous(name="Days post-ischemia (DPI) ",
                     breaks=c(14,30)) +
  Plot_theme
  
Striatum_Pdgfr_Low_Sctr | Striatum_Pdgfr_High_Sctr
```

@Pdgfrb_Striatum_LowHigh_Exploratory show that PDGFR-Î²^low^ (non-reactive cells) show a wide dispersion give the reduced number of observations at 14 DPI. Meanwhile, PDGFR-Î²^high^ display an increasing trend from 14 to 30 DPI. As done for brain with cortico-striatal lesions, we fit the model using a binomial family distribution. Please refer to the `Widefield_5x_Pdgfr-LowHigh/qmd` notebook for additional details of this dristribution model. 

# Statistical modeling 

## Fit the model

We fit a linear model using a binomial distribution `DPI` as a linear predictor for the probability of PDGFR-ð›½^high^:

$$
\log\left(\frac{p_{i}}{1 - p_{i}}\right) = \alpha + \beta_{1} \times DPI_{i}
$$ 
This model uses weakly informative priors based on the information gathered from the model for corticostriatal lesions. The prior take the following notation:  
 
$$
\begin{align}
b \sim \mathcal{N}(1, 0.2),  \text{lb}=0 \\
\text{Intercept} \sim student_t(3, 0, 2), \text{lb}=0 \\
\end{align}
$$

```{r}
#| label: Pdgfrb_LowHigh_Str_Modeling
#| include: true
#| warning: false
#| message: false
#| results: false
#| cache: true

Pdgfr_Prop_Str_Mdl1 <- bf(Pdgfr_React | trials(Pdgfr_Total) ~ DPI)

get_prior(Pdgfr_Prop_Str_Mdl1, Pdgfrb_Striatum, family = binomial())

Pdgfr_Prop_Str_Prior1 <- 
  c(prior(normal(1,0.2), class = b, lb = 0),
    prior(student_t(3, 0, 2), class = Intercept, lb = 0))

# Fit model 1
Pdgfr_Prop_Str_Fit1 <- 
  brm(
    data    = Pdgfrb_Striatum,
    family  = binomial(), 
    formula = Pdgfr_Prop_Str_Mdl1,
    prior   = Pdgfr_Prop_Str_Prior1,
    chains  = 4,
    cores   = 4,
    warmup  = 2500, 
    iter    = 5000, 
    seed    = 8807,
    control = list(adapt_delta = 0.99, max_treedepth = 15),
    file    = "Models/Pdgfrb_LowHigh/Pdgfrb_Prop_Str_Fit1.rds",
    file_refit = "never") 
                     
# Add loo for model comparison
Pdgfr_Prop_Strl_Fit1 <- 
  add_criterion(Pdgfr_Prop_Str_Fit1, c("loo", "waic", "bayes_R2"))
```

## Model diagnostics

We perform model checks using the `pp_check` function:

```{r}
#| label: Pdgfrb_LowHigh_Str_Diagnistics 
#| include: true
#| warning: false
#| message: false
#| results: false
#| cache: true
#| fig-cap: Model diagnostics for the probability of PDGFR-ð›½^high^ (striatal lesions)
#| fig-height: 4
#| fig-width: 5

set.seed(8807)

Pdgfr_Prop_Str_Mdl1_pp <- 
  brms::pp_check(Pdgfr_Prop_Str_Fit1, 
                 ndraws = 100) +
    labs(title = expression("Posterior predictive checks (PDGFR-Î²)"),
    subtitle = "Formula: PDGFR_React | PDGFR_Total ~ DPI") +
  Plot_theme  
 
Pdgfr_Prop_Str_Mdl1_pp
```

Similar to our cortico-striatal injuries, our predictions exhibit a moderate deviations from the data. We further diagnose calculating Leave-One-Out Cross-Validation (Loo) using the `loo` function from `brms`.

```{r}
#| label: Pdgfrb_Str_Loo
#| include: true
#| warning: false
#| message: false

loo(Pdgfr_Prop_Str_Fit1)
```

The results show that most observations are problematic. In particular, 48% of observations belonging to the very bad category, suggest that the importance sampling weights for these observations are highly variable. Therefore, it is fair to ponder that this model must be considered with caution.

# Model results

## Visualization of conditional effects

```{r}
#| label: fig-Pdgfr_LowHigh_Str_CondEff
#| include: true
#| warning: false
#| message: false
#| results: false
#| fig-cap: Posterior for PDGFR-ð›½^low^ and PDGFR-ð›½^high^ (striatal lesions)
#| fig-width: 4
#| fig-height: 3

set.seed(8807)

# We create the graph for convex hull
Pdgfr_Prop_Str_DPI <- 
  conditional_effects(Pdgfr_Prop_Str_Fit1)

Pdgfr_Prop_Str_DPI <- plot(Pdgfr_Prop_Str_DPI, 
       plot = FALSE)[[1]]

Pdgfr_Prop_Str_fig <- Pdgfr_Prop_Str_DPI  + 
  scale_y_continuous(name = expression ("(p) PDGFR-Î²"^high)) +
  scale_x_continuous(name="DPI") +
  scale_color_manual(
    values = c("#0048BA", "red"),
    name="Condition"
    ) +
  scale_fill_manual(
    values = c("#0048BA", "red"),
    name="Condition"
    ) +

  Plot_theme +
  theme(legend.position = "top", legend.direction = "horizontal")

ggsave(
  plot     = Pdgfr_Prop_Str_fig, 
  filename = "Plots/Pdgfr_Prop_Str_fig.png", 
  width    = 9, 
  height   = 9, 
  units    = "cm")

Pdgfr_Prop_Str_fig
```

@fig-PDGFRProp_Striatal_CondEff show an increasing probability for reactive PDGFR-Î² with a low uncertainty. This shows a clear contrast with cortico-striatal lesions which exhibit twice the probability at 14 and 30 DPI.

## Posterior summary

Here, we visualize the results in numerical terms using the `describe_posterior` function:

```{r}
#| label: Pdgfr_LowHigh_Str_DescribePosterior
#| include: true
#| warning: false
#| message: false
#| results: false

describe_posterior(
  Pdgfr_Prop_Str_Fit1,
  effects = "all",
  test = c("p_direction", "rope"),
  component = "all",
  centrality = "median")
```
