---
title-block-banner: true
title: "Analysis of PDGFR-Î²-low and PDGFR-Î² high+ cells in the ipsilateral hemisphere"
subtitle: "Data analysis notebook"
date: today
date-format: full
author: 
  - name: "Daniel Manrique-Castano"
    orcid: 0000-0002-1912-1764
    degrees:
      - PhD
    affiliation: 
      - name: Univerisity Laval 
        department: Psychiatry and Neuroscience
        group: Laboratory of neurovascular interactions 
note: "GitHub: https://daniel-manrique.github.io/"

keywords: 
  - PDGFR-Î²
  - Brain injury
  - Bayesian modeling 
   
license: "CC BY"

format:
   pdf: 
    toc: true
    number-sections: true
    colorlinks: true
   html:
    code-fold: true
    embed-resources: true
    toc: true
    toc-depth: 2
    toc-location: left
    theme: spacelab

knitr:
  opts_chunk: 
    warning: false
    message: false

csl: science.csl
bibliography: references.bib
---

# Preview

In this notebook, we analyze the proportion and distribution of PDGFR-Î²_Low and PDGFR-Î²_High+ cells with automatic cell detection and classification using QuPath [@bankhead2017]. We also create point patterns using the coordinates of detecting cells to perform Point Pattern analysis. 

**Parent dataset:** PDGFR-Î² and GFAP-stained ischemic hemispheres imaged at 5x (with stitching). Samples are grouped at 0 (Sham), 3, 7, 14, and 30 days post-ischemia (DPI). The raw images and pre-processing scripts (if applicable) are available at the Zenodo repository (10.5281/zenodo.10553084) under the name `Widefield_5x_Ipsilateral_Gfap-Pdgfrb.zip`. Individual cells were detected and classified into PDGFR-Î²^low^ (Pdgfrb_NonReact) and PDGFR-Î²^high^ (Pdgfrb_React) using QuPath [@bankhead2017].The complete QuPath project, including classifiers and output data as .tsv files is available at https://osf.io/8ehyu.

**Working dataset**: The `Data_Processed/Widefield_5x_Ipsilateral_Gfap-Pdgfrb/Widefield_5x_Ipsilateral_Gfap-Pdgfrb_Inten.csv`data frame containing the number of PDGFR-Î²^low^ (Pdgfrb_NonReact) and PDGFR-Î²^high^ (Pdgfrb_React) cells in the ischemic hemisphere. Here, we analyze the proportion and distribution of these populations.

# Install and load required packages

Install and load all required packages. Please uncomment (delete #) the line code if installation is required. Load the installed libraries each time you start a new R session.

```{r}
#| label: Install_Packages
#| include: true
#| warning: false
#| message: false

#install.packages("devtools")
#library(devtools)

#install.packages(c("bayesplot", "bayestestR", "brms","broom.mixed", "dplyr", "easystats", "distributional", "ggplot","gtsummary", "modelbased", "modelr", "modelsummary", "patchwork", "poorman", "tidybayes", "tidyverse"))

library(bayesplot)
library(bayestestR)
library(brms)
library(broom.mixed)
library(dplyr)
library(easystats)
library(distributional)
library(ggplot2)
library(gtsummary)
library(modelbased)
library(modelr)
library(modelsummary)
library(patchwork)
library(poorman)
library(tidybayes)
library(tidyverse)
```

# Visual themes

We create a visual theme to use in our plots (ggplots).

```{r}
#| label: Plot_Theme
#| include: true
#| warning: false
#| message: false
  
Plot_theme <- theme_classic() +
  theme(
      plot.title = element_text(size=18, hjust = 0.5, face="bold"),
      plot.subtitle = element_text(size = 10, color = "black"),
      plot.caption = element_text(size = 12, color = "black"),
      axis.line = element_line(colour = "black", size = 1.5, linetype = "solid"),
      axis.ticks.length=unit(7,"pt"),
     
      axis.title.x = element_text(colour = "black", size = 16),
      axis.text.x = element_text(colour = "black", size = 16, angle = 0, hjust = 0.5),
      axis.ticks.x = element_line(colour = "black", size = 1),
      
      axis.title.y = element_text(colour = "black", size = 16),
      axis.text.y = element_text(colour = "black", size = 16),
      axis.ticks.y = element_line(colour = "black", size = 1),
      
      legend.position="right",
      legend.direction="vertical",
      legend.title = element_text(colour="black", face="bold", size=12),
      legend.text = element_text(colour="black", size=10),
      
      plot.margin = margin(t = 10,  # Top margin
                             r = 2,  # Right margin
                             b = 10,  # Bottom margin
                             l = 10) # Left margin
      ) 
```


# Exploratory data visualization

We load the `Data_Processed/Widefield_5x_Ipsilateral_Gfap-Pdgfrb/Widefield_5x_Ipsilateral_Gfap-Pdgfrb_Inten.csv` dataset to very its content.

```{r}
#| label: tbl-Pdgfrb_LowHigh_Table
#| include: true
#| warning: false
#| message: false
#| tbl-cap: "Data set"

# We load the dataset in case is not present in the R environment
Pdgfrb_Summary <- read.csv(file = "Data_Processed/Widefield_5x_Ipsilateral_Gfap-Pdgfrb/Widefield_5x_Ipsilateral_Gfap-Pdgfrb_Inten.csv", header = TRUE)

gt::gt(Pdgfrb_Summary[1:10,])
```

From this table, we focus on `DPI` (Days post-ischemia), `Pdgfrb_Neg`, `Pdgfrb_Pos` variables to analyze these cells proportions in the ischemic brain. Next, we visualize the raw data to guide the statistical modeling. We plot the response variables as a density and a scatter plot (per DPI). In the scatter plot, we fit lines for a lineal (black), 2-degree (red), and 3-degree (green) polynomial models.

```{r}
#| label: fig-Pdgfrb_LowHigh_Exploratory
#| include: true
#| warning: false
#| message: false
#| results: false
#| fig-cap: Exploratory data visualization for PDGFR-Î²^low^ and PDGFR-Î²^high^
#| fig-width: 9
#| fig-height: 4

set.seed(8807)

# PDGFR-Î²^low^
##################

Pdgfrb_Low_Sctr <- 
  ggplot(
    data  = Pdgfrb_Summary, 
    aes(x = DPI, 
        y = Pdgfrb_Neg)) +
geom_smooth(
  method = "lm", 
  se     = TRUE,
  color  = "black") +
geom_smooth(
  method  = "lm", 
  se      = TRUE, 
  formula = y ~ poly(x, 2), 
  color   = "darkred") +
geom_smooth(
  method  = "lm", 
  se      = TRUE, 
  formula = y ~ poly(x, 3), 
  color   = "darkgreen") +
geom_jitter(
  width = 0.5, 
  shape = 1, 
  size  = 1.5, 
  color = "black") +
  scale_y_continuous(name= expression("Number of PDGFR-Î²"^low)) +
  scale_x_continuous(name="DPI",
                     breaks=c(0, 3, 7,14,30)) +
  Plot_theme
  

# PDGFR-Î²^high^
######################

Pdgfrb_High_Sctr <- 
  ggplot(
    data  = Pdgfrb_Summary, 
    aes(x = DPI, 
        y = Pdgfrb_Pos)) +
geom_smooth(
  method = "lm", 
  se     = TRUE,
  color  = "black") +
geom_smooth(
  method  = "lm", 
  se      = TRUE, 
  formula = y ~ poly(x, 2), 
  color   = "darkred") +
geom_smooth(
  method  = "lm", 
  se      = TRUE, 
  formula = y ~ poly(x, 3), 
  color   = "darkgreen") +
geom_jitter(
  width = 0.5, 
  shape = 1, 
  size  = 1.5, 
  color = "black") +
  scale_y_continuous(name= expression("Number of PDGFR-Î²"^high)) +
  scale_x_continuous(name="DPI",
                     breaks=c(0, 3, 7,14,30)) +
  Plot_theme


Pdgfrb_Low_Sctr | Pdgfrb_High_Sctr
```

@PdgfrbLowHigh_Exploratory show that PDGFR-Î²^low^ (non-reactive cells) do not fit well due to the sharp drop at 3 DPI. On the other hand, we see that non-linear models are a better alternative for PDGFR-Î²^high^ cells. AS expected, the precedent implies that the reactivity patterns observed previously are mostly mediated by PDGFR-Î²^high^ (reactive) cells. We take this into consideration for our modeling.

# Statistical modeling for the proportion of PDGFR-Î²^low^ and PDGFR-Î²^high^ cells

Considering that PDGFR-Î²^low^ and PDGFR-Î²^high^ cells are mirror populations conditional on the total number of PDGFR-Î² cells, we will use a model to analyze the cell proportions. For this purpose, we employ the binomial family distribution, where the response variable represents a series of Bernoulli trials (PDGFR-Î²^high^ or PDGFR-Î²^low^)in a fixed number of independent trials (PDGFR-ð›½^total^). This family is particularly well-suited for interpreting the underlying event probabilities.

Mathematically, the probability mass function (PMF) for a binomial distribution is given as:

$$
P(y | n, p) = \binom{n}{y} p^y (1 - p)^{n - y}
$$ 

Where: - $y$ is the number of successes. - $n$ is the number of trials. - $p$ is the probability of success on an individual trial. - $\binom{n}{y}$ is the binomial coefficient, representing the number of ways to choose $y$ successes in $n$ trials. In `brms`, the linear predictor $\eta$ is linked to the probability $p$ of success using a the logit function:

$$
log\left(\frac{p}{1 - p}\right) = \eta
$$ 

We fit the following models:

-   **Pdgfrb_Prop_Mdl1:** We use `DPI` as a linear predictor for the probability of PDGFR-ð›½^high^:

$$
\log\left(\frac{p_{i}}{1 - p_{i}}\right) = \alpha + \beta_{1} \times DPI_{i}
$$ 

Where: - $p_{i}$ is the probability of `Pdgfr_React` being a success on the $i^{th}$ trial and \$DPI\_{i} is the $i^{th}$ observed value of `DPI`.

-   **Pdgfrb_Prop_Mdl2:** We use `DPI` with splines and 5 knots:

$$
\log\left(\frac{p_{i}}{1 - p_{i}}\right) = \alpha + s(DPI_{i}, k = 5)
$$ 

Where: - $p_{i}$ is the probability of `Pdgfrb_React` being a success on the $i^{th}$ trial. - $DPI_{i}$ is the $i^{th}$ observed value of `DPI`. - $s(DPI_{i}, k = 5)$ is a smooth function of `DPI` with 5 basis functions.

The observed counts for `Pdgfrb_React` out of `Pdgfrb_Total` are modeled as a binomial distribution with probability $p_{i}$. In both cases, we use flat-default `brms` priors.

## Fit the models

```{r}
#| label: Pdgfrb_LowHigh_Modeling
#| include: true
#| warning: false
#| message: false
#| results: false
#| cache: true


# Model 1: DPI as a linear predictor
########################################

Pdgfrb_Prop_Mdl1 <- bf(Pdgfrb_Pos | trials(Pdgfrb_Total) ~ DPI)

get_prior(Pdgfrb_Prop_Mdl1, Pdgfrb_Summary, family = binomial())

# Fit model 1
Pdgfrb_Prop_Fit1 <- 
  brm(
    data    = Pdgfrb_Summary,
    family  = binomial(), 
    formula = Pdgfrb_Prop_Mdl1,
    chains  = 4,
    cores   = 4,
    warmup  = 2500, 
    iter    = 5000, 
    seed    = 8807,
    control = list(adapt_delta = 0.99, max_treedepth = 15),
    file    = "Models/Widefield_5x_Ipsilateral_Pdgfrb_LowHigh/Widefield_5x_Ipsilateral_LowHigh_Fit1.rds",
    file_refit = "never") 
                     
# Add loo for model comparison
Pdgfrb_Prop_Fit1 <- 
  add_criterion(Pdgfrb_Prop_Fit1, c("loo", "waic", "bayes_R2"))


# Model 1: DPI as predictor with splines
#############################################

Pdgfrb_Prop_Mdl2 <- bf(Pdgfrb_Pos | trials(Pdgfrb_Total) ~ s(DPI, k =5))

get_prior(Pdgfrb_Prop_Mdl2, Pdgfrb_Summary, family = binomial())

# Fit model 2
Pdgfrb_Prop_Fit2 <- 
  brm(
    data    = Pdgfrb_Summary,
    family  = binomial(), 
    formula = Pdgfrb_Prop_Mdl2,
    knots   = list(DPI = c(0, 3, 7, 14, 30)),
    chains  = 4,
    cores   = 4,
    warmup  = 2500, 
    iter    = 5000, 
    seed    = 8807,
    control = list(adapt_delta = 0.99, max_treedepth = 15),
    file    = "Models/Widefield_5x_Ipsilateral_Pdgfrb_LowHigh/Widefield_5x_Ipsilateral_LowHigh_Fit2.rds",
    file_refit = "never") 
                     
# Add loo for model comparison
Pdgfrb_Prop_Fit2 <- 
  add_criterion(Pdgfrb_Prop_Fit2, c("loo", "waic", "bayes_R2"))
```

## Model comparison

We perform model comparison using the WAIC criteria. Please refer to `Widefield_5x_Ipsilateral_Pdgfrb_IntDen` notebook for further details on this procedure. 

```{r}
#| label: Pdgfrb_LowHigh_Compare
#| include: true
#| warning: false
#| message: false
#| results: false

Pdgfrb_Prop_Comp <- 
  compare_performance(
    Pdgfrb_Prop_Fit1, 
    Pdgfrb_Prop_Fit2 
    )

Pdgfrb_Prop_Comp
```
In both models, R2 is over 0.9. However, we can see that model 2 (with splines) is far less penalized for out of sample prediction (5084 vs 12480). We visualize the same results as a graph:

```{r}
#| label: fig-Pdgfrb_LowHigh_Compare
#| include: true
#| warning: false
#| message: false
#| results: false
#| fig-cap: Model camparison by WAIC
#| fig-height: 4
#| fig-width: 5

Pdgfrb_Prop_W <- 
loo_compare(
  Pdgfrb_Prop_Fit1, 
  Pdgfrb_Prop_Fit2, 
  criterion = "waic")

# Generate WAIC graph
Pdgfrb_Prop_WAIC <- 
  Pdgfrb_Prop_W[, 7:8] %>% 
  data.frame() %>% 
  rownames_to_column(var = "model_name") %>% 
  
ggplot(
  aes(x    = model_name, 
      y    = waic, 
      ymin = waic - se_waic, 
      ymax = waic + se_waic)
  ) +
  geom_pointrange(shape = 21) +
  scale_x_discrete(
    breaks=c("Pdgfrb_Prop_Fit1", 
             "Pdgfrb_Prop_Fit2"), 
             
    labels=c("Mdl1", 
             "Mdl2")) +
    
  coord_flip() +
  labs(x = "", 
       y = "WAIC (score)",
       title = "") +
  Plot_theme

Pdgfrb_Prop_WAIC
```

We have sufficient grounds to continue with model 2 for scientific inference.

## Model diagnostics

We check the model fitting using posterior predictive checks

```{r}
#| label: fig-PdgfrbProp_Diagnistics
#| include: true
#| warning: false
#| message: false
#| results: false
#| fig-cap: Model diagnostics for the probability of PDGFR-Î²^high^
#| fig-height: 4
#| fig-width: 5

set.seed(8807)

Pdgfrb_Prop_Mdl2_pp <- 
  brms::pp_check(Pdgfrb_Prop_Fit2, 
                 ndraws = 100) +
  labs(title = expression("Posterior predictive checks (Pdgfrb-Î²)"),
  subtitle = "Formula: Pdgfrb_React | Pdgfrb_Total ~ s(DPI, k = 5)") +
  Plot_theme 

Pdgfrb_Prop_Mdl2_pp
```
We see that the predictions do not match accurately the data, but follow the same trend. No consider this does not constitute a significant deviation but results must be addressed with caution. We can explore further the model using `shinystan`.

```{r}
#| label: Pdgfrb_LowHigh_Shiny
#| include: true
#| warning: false
#| message: false
#| results: false
#| cache: true

#launch_shinystan(Pdgfrb_Prop_Fit2)
```

# Model results

## Visualization of conditional effects

We use the `conditional_effects` function from `brms` to visualize the results:

```{r}
#| label: fig-Pdgfrb_LowHigh_CondEff
#| include: true
#| warning: false
#| message: false
#| results: false
#| fig-cap: Posterior for PDGFR-Î²^high^
#| fig-width: 5
#| fig-height: 5

set.seed(8807)

# We create the graph for convex hull
Pdgfrb_Prop_DPI <- 
  conditional_effects(Pdgfrb_Prop_Fit2)

Pdgfrb_Prop_DPI <- plot(Pdgfrb_Prop_DPI, 
       plot = FALSE)[[1]]

Pdgfrb_Prop_fig <- Pdgfrb_Prop_DPI  + 
  scale_y_continuous(name = expression ("(p) PDGFR-Î²"^high)) +
  scale_x_continuous(name="DPI") +
  scale_color_manual(
    values = c("#0048BA", "red"),
    name="Condition"
    ) +
  scale_fill_manual(
    values = c("#0048BA", "red"),
    name="Condition"
    ) +

  Plot_theme +
  theme(legend.position = "top", legend.direction = "horizontal")

ggsave(
  plot     = Pdgfrb_Prop_fig, 
  filename = "Plots/Widefield_5x_Ipsilateral_Pdgfrb_LowHigh/Widfield_5x_Ipsilateral_Pdgfrb_LowHigh.png", 
  width    = 9, 
  height   = 9, 
  units    = "cm")

Pdgfrb_Prop_fig
```

@fig-Pdgfrb_LowHigh_CondEff show an increasing probability for reactive PDGFR-Î². This is largely consistent with the integrated density measurements, suggesting that PDGFR-Î² reactivity in chronic stages is largely driven by PDGFR-Î²^high^ cells.

## Posterior summary

We plot the posterior summary using the `describe_posterior` function:

```{r}
#| label: Pdgfrb_LowHigh_DescribePosterior
#| include: true
#| warning: false
#| message: false
#| results: false
#| cache: true

describe_posterior(
  Pdgfrb_Prop_Fit2,
  effects = "all",
  test = c("p_direction", "rope"),
  component = "all",
  centrality = "median")

modelsummary(Pdgfrb_Prop_Fit2, 
             shape = term ~ model + statistic,
             centrali2ty = "mean", 
             title = "PDGFR-Î² low and high-intensity populations following MCAO",
             statistic = "conf.int",
             gof_omit = 'ELPD|ELDP s.e|LOOIC|LOOIC s.e|WAIC|RMSE',
             output = "Tables/html/Widefield_5x_Ipsilateral_Pdgfrb_LowHigh_Fit2_Table.html",
             )

Pdgfrb_LowHigh_Fit2_Table <- modelsummary(Pdgfrb_Prop_Fit2, 
             shape = term ~ model + statistic,
             centrality = "mean", 
             statistic = "conf.int",
             gof_omit = 'ELPD|ELDP s.e|LOOIC|LOOIC s.e|WAIC|RMSE',
             output = "gt")
gt::gtsave (Pdgfrb_LowHigh_Fit2_Table, filename = "Tables/tex/Widefield_5x_Ipsilateral_Pdgfrb_LowHigh_Fit2_Table.tex")
```

We did not found a tool the calculate derivatives from binomial models. Therefore, we must perform scientific inference based on the provided conditional effects. We can visualize a sharp increase in the probability of PDGFR-Î²^high^ up to the second-third weeks post-ischemia. Followed by the plateau phase indicated by the integrated density estimations.


::: {#refs}
:::

```{r}
sessionInfo()
```
