---
title-block-banner: true
title: "Analysis of Klf4 expression in the ipsilateral hemisphere"
subtitle: "Data analysis notebook"
date: today
date-format: full
author: 
  - name: "Daniel Manrique-Castano"
    orcid: 0000-0002-1912-1764
    affiliation: Univerisity Laval (Laboratory of neurovascular interactions) 
keywords: 
  - Klf4
  - Brain injury
  - Bayesian modeling 
   
license: "CC BY"

format:
   pdf: 
    toc: true
    number-sections: true
    colorlinks: true
   html:
    code-fold: true
    embed-resources: true
    toc: true
    toc-depth: 2
    toc-location: left
    number-sections: true
    theme: spacelab

knitr:
  opts_chunk: 
    warning: false
    message: false
    
csl: science.csl
bibliography: References.bib
---

# Preview

This notebook reports the analysis of KLF4 expression in the ipsilateral hemisphere following cerebral ischemia.

**Parent dataset:** KLF4 and PDGFR-Î² stained ischemic hemispheres imaged at 10x (with stitching). Samples are grouped at 0 (Sham), 3, 7, 14, and 30 days post-ischemia (DPI). The raw images and pre-processing scripts (if applicable) are available at the Zenodo repository (10.5281/zenodo.10553084) under the name `Widefield_10x_Ipsilateral_Klf4.zip`.

**Working dataset**: The `Data_Raw/Widefield_10x_Klf4_Detections/Widefield_10x_Ipsilateral_Klf4_Cells.csv`and `Data_Raw/Widefield_10x_Klf4_Detections/Widefield_10x_Ipsilateral_Klf4_Image.csv`data frames containing the raw output from CellProfiller [@stirling2021]. The CellProfiller pipeline used to perform the KLF4+ cell detection is available at https://osf.io/4fh32.

We perform scientific inference based on the number, intensity and distribution of KLF4+ cells. We anticipate that KLF4 is induced after ischemia.

# Install and load required packages

Install and load all required packages. Please uncomment (delete #) the line code if installation is required. Load the installed libraries each time you start a new R session.

```{r}
#| label: Install_Packages
#| include: true
#| warning: false
#| message: false

#install.packages("devtools")
library(devtools)

#install.packages(c("bayesplot", "bayestestR", "brms","dplyr", "easystats", "ggplot","modelbased", "modelr", "patchwork", "poorman","plyr", "spatstat", "tidybayes", "tidyverse", "viridis"))


library(bayesplot)
library(bayestestR)
library(brms)
library(dplyr)
library(easystats)
library(emmeans)
library(ggplot2)
library(modelbased)
library(modelr)
library(patchwork)
library(poorman)
library(plyr)
library(spatstat)
library(tidybayes)
library(tidyverse)
library(viridis)
```

# Visual themes

We create a visual theme to use in our plots (ggplots).

```{r}
#| label: Plot_Theme
#| include: true
#| warning: false
#| message: false
  
Plot_theme <- theme_classic() +
  theme(
      plot.title = element_text(size=18, hjust = 0.5, face="bold"),
      plot.subtitle = element_text(size = 10, color = "black"),
      plot.caption = element_text(size = 12, color = "black"),
      axis.line = element_line(colour = "black", size = 1.5, linetype = "solid"),
      axis.ticks.length=unit(7,"pt"),
     
      axis.title.x = element_text(colour = "black", size = 16),
      axis.text.x = element_text(colour = "black", size = 16, angle = 0, hjust = 0.5),
      axis.ticks.x = element_line(colour = "black", size = 1),
      
      axis.title.y = element_text(colour = "black", size = 16),
      axis.text.y = element_text(colour = "black", size = 16),
      axis.ticks.y = element_line(colour = "black", size = 1),
      
      legend.position="right",
      legend.direction="vertical",
      legend.title = element_text(colour="black", face="bold", size=12),
      legend.text = element_text(colour="black", size=10),
      
      plot.margin = margin(t = 10,  # Top margin
                             r = 2,  # Right margin
                             b = 10,  # Bottom margin
                             l = 10) # Left margin
      ) 
```

# Analysis of KLF4+ spatial inensity

## Load and handle the datasets

We load the `Data_Raw/Widefield_10x_Klf4_Detections/Widefield_10x_Ipsilateral_Klf4_Cells.csv` and `Data_Raw/Widefield_10x_Klf4_Detections/Widefield_10x_Ipsilateral_Dapi_Cells.csv` datasets. These are heavy data sets containing information about individual cells detected in the ipsilateral hemisphere.

```{r}
#| label: tbl-Klf4_Table
#| include: true
#| warning: false
#| message: false
#| tbl-cap: "Data set"

# We load the dataset in case is not present in the R environment
Klf4_Cells <- read.csv(file = "Data_Raw/Widefield_10x_Klf4_Detections/Widefield_10x_Ipsilateral_Klf4_Cells.csv", header = TRUE)

Dapi_Cells <- read.csv(file = "Data_Raw/Widefield_10x_Klf4_Detections/Widefield_10x_Ipsilateral_Dapi_Cells.csv", header = TRUE)

knitr::kable(Klf4_Cells [1:10, ])
```

From the KLF4 table, we are interested in the `FileName_RawI`column containing the identification data for the images, the `Intensity_MedianIntensity_Raw` indicating the median intensity of individual KLF4 detection, and `Location_Center_X` and `Location_Center_Y`signaling the xy coordinates of each detection. From the Dapi table, we are interested in the metadta and the coordinates to set up the observation window. Next, we subset the dataset to select the columns of interest and give them meaningful names.

```{r}
#| label: tbl-Klf4_Handle
#| include: true
#| warning: false
#| message: false
#| tbl-cap: "Data set"

## We subset the relevant columns (cell number)
Klf4_Data <- subset(Klf4_Cells, select = c("FileName_Raw", "Intensity_MedianIntensity_Raw", "Location_Center_X", "Location_Center_Y"))

## And extract metadata from the image name
Klf4_Data  <- cbind(Klf4_Data, do.call(rbind , strsplit(Klf4_Data$FileName_Raw, "[_\\.]"))[,1:2])

Klf4_Data <- subset(Klf4_Data, select = -c(FileName_Raw))

## We Rename the relevant columns 
colnames(Klf4_Data) <- c("Intensity", "CenterX", "CenterY", "MouseID", "DPI")

## We set the factors
Klf4_Data$DPI <- factor(Klf4_Data$DPI, levels = c("0D", "3D", "7D", "14D", "30D"))

write.csv(Klf4_Data, "Data_Processed/Wide10x_Ipsilateral_Klf4_CellIntensity.csv", row.names = FALSE)


## For DAPI

## We subset the relevant columns (cell number)
Dapi_Data <- subset(Dapi_Cells, select = c("FileName_Raw", "Intensity_MedianIntensity_Raw", "Location_Center_X", "Location_Center_Y"))

## And extract metadata from the image name
Dapi_Data  <- cbind(Dapi_Data, do.call(rbind , strsplit(Dapi_Data$FileName_Raw, "[_\\.]"))[,1:2])

Dapi_Data <- subset(Dapi_Data, select = -c(FileName_Raw))

## We Rename the relevant columns 
colnames(Dapi_Data) <- c("Intensity", "CenterX", "CenterY", "MouseID", "DPI")

## We set the factors
Dapi_Data$DPI <- factor(Dapi_Data$DPI, levels = c("0D", "3D", "7D", "14D", "30D"))

write.csv(Dapi_Data, "Data_Processed/Wide10x_Ipsilateral_Dapi_CellIntensity.csv", row.names = FALSE)

```

With the data handled, we create point patterns to perform Point Pattern Analysis (PPA). The advantage of this technique is that it takes into account the different hemispheric areas to estimate the spatial intensity of cells. For more information on PPA, please refer to the `Widefield_10x_Pdgfrb-Gfap_Covariance`notebook.

## Generate and handle point patterns

We use functions from the \`spatstat' package to create point patterns based on the coordinates of individual cells. The point patterns are then stored in a hyperframe and can be loaded into R as an R object.

```{r}
#| label: tbl-Klf4_PointPatterns
#| include: true
#| warning: false
#| message: false
#| tbl-cap: "Data set"

# Initialize the hyperframe as NULL at the start
Result_Hyperframe <- NULL

# Adjusted add_to_hyperframe function to dynamically build the hyperframe
add_to_hyperframe <- function(...) {
  if (is.null(Result_Hyperframe)) {
    Result_Hyperframe <<- hyperframe(...)
  } else {
    Result_Hyperframe <<- tryCatch({
      rbind(Result_Hyperframe, hyperframe(...))
    }, error = function(e) {
      cat("Error in rbind: ", e$message, "\n")
    })
  }
}

# Adjusted function to create point patterns
create_point_pattern <- function(Data_Subset) {
  xlim <- range(Data_Subset$CenterX)
  ylim <- range(Data_Subset$CenterY)

  Cells_PPP <- spatstat.geom::ppp(x = Data_Subset$CenterX, y = Data_Subset$CenterY, xrange = xlim, yrange = ylim)
  unitname(Cells_PPP) <- list("mm", "mm", 3.007/6624)
  Cells_PPP <- spatstat.geom::rescale(Cells_PPP)
  
  return(Cells_PPP)
}

# Iterate over unique MouseIDs to process and create point patterns for both Klf4 and Dapi
mouse_ids <- unique(Klf4_Data$MouseID)

for (mouse_id in mouse_ids) {
  Subset_Klf4 <- Klf4_Data[Klf4_Data$MouseID == mouse_id, ]
  Subset_Dapi <- Dapi_Data[Dapi_Data$MouseID == mouse_id, ]
  
  if(nrow(Subset_Klf4) > 0 && nrow(Subset_Dapi) > 0) {
    Klf4_PPP <- create_point_pattern(Subset_Klf4)
    Dapi_PPP <- create_point_pattern(Subset_Dapi)
    
    # Set the observation window for Klf4 based on Dapi's convex hull
    Window(Klf4_PPP) <- convexhull(Dapi_PPP)
    
    dpi2_value <- unique(Subset_Klf4$DPI)[1]
    
    add_to_hyperframe(Klf4 = Klf4_PPP, Dapi = Dapi_PPP, ID = as.character(mouse_id), DPI = as.factor(dpi2_value), stringsAsFactors = TRUE)
  } else {
    message(sprintf("Skipping MouseID %s due to insufficient data.\n", mouse_id))
  }
}

# Save the Result_Hyperframe
Result_Hyperframe$DPI <- factor(Result_Hyperframe$DPI, levels = c("0D", "3D", "7D", "14D", "30D"))

saveRDS(Result_Hyperframe, "PointPatterns/Widefield_10x_Ipsilateral_Klf4.rds")
```

The point patterns are created and stored. In the next, chunk, we load the point patterns and add density kernels. Please check the `Widefield_10x_Pdgfrb-Gfap_Covariance`notebook for more information in this regard.

## Calculate density kernels

We use the `density` function to calculate density kernels with a sigma of 0.02.

```{r}
#| label: tbl-Klf4_Densitykernels
#| include: true
#| warning: false
#| message: false

# Load the point patterns
PointPatterns <- readRDS("PointPatterns/Widefield_10x_Ipsilateral_Klf4.rds")

# Add density kernels to the hyperframe
PointPatterns$Klf4_Density <- with (PointPatterns, density(Klf4, sigma = 0.02))
```

### Plot density kernels

For visualization purposes, we plot some examples of the density kernels. Please not that the coordinates are y-flied \*for unknown reason to us.This, however, does not hinder the main message, that KLF4+ cells tend to aggregate in the ischemic cortex (to the left) as the injury progresses.

```{r}
#| label: tbl-Klf4_Plotkernels
#| include: true
#| warning: false
#| message: false
#| fig-cap: Example density kernels for KLF4
#| fig-height: 5
#| fig-width: 9

Klf4_Colmap <- colourmap(topo.colors(256), range = c(0, 20000))

par(mfrow = c(2,5), mar=c(1,1,1,1), oma=c(1,1,1,1))

plot(PointPatterns$Klf4_Density$`138`, col = Klf4_Colmap, main = "0 DPI")  
plot(PointPatterns$Klf4_Density$`140`, col = Klf4_Colmap, main = "3 DPI")
plot(PointPatterns$Klf4_Density$`13`, col = Klf4_Colmap, main = "7 DPI")
plot(PointPatterns$Klf4_Density$`125`, col = Klf4_Colmap, main = "14 DPI")
plot(PointPatterns$Klf4_Density$`117`, col = Klf4_Colmap, main = "30 DPI")
```
## Modeling point pattern process

In this section we use the `mppm` function from `spatstat` to fit a **loglinea models** for replicated point patterns. This model allow us to quantify KLF4 allocation changes conditional on the x coordinate of the ischemic hemisphere. This is relevant for us given the specific injury site. In this case, we use a multilevel model with random intercepts and slopes for the x coordinate. The model specification is as follows: 

## Statistical Model Specification

The spatial intensity of KLF4 expression was modeled using a multiplicative Poisson point process model (mppm), applied to a dataset comprising 52 point patterns. The model aimed to assess how the spatial distribution of KLF4 expression varies with the x-coordinate, conditional on the developmental period (DPI). The model can be specified in statistical notation as follows:

$$
KLF4_i(x) \sim \text{Poisson}(\lambda(x, \text{DPI}_i)) 
$$

where $i$ indexes the point pattern (corresponding to a unique DPI), and $x$ represents the spatial coordinate. The intensity function $\lambda(x, \text{DPI}_i)$ is modeled as:

$$
\log(\lambda(x, \text{DPI}_i)) = \beta_0 + \beta_1 x + u_{0i} + u_{1i}x
$$
Where $\beta_0$ and $\beta_1$ are fixed effects, representing the baseline log intensity of KLF4 expression and the effect of the x-coordinate on this intensity, respectively.$u_{0i}$ and $u_{1i}$ are random effects for the intercept and slope, varying by DPI, to capture the variability in KLF4 expression intensity and its spatial variation across different injury stages.

### Fit the mppm model

We fit the model using the `mppm` function for treating replicated point patterns as those in out hyperframe. 

```{r}
#| label: tbl-Klf4_mppm
#| include: true
#| warning: false
#| message: false
#| fig-cap: Example density kernels for KLF4
#| fig-height: 5
#| fig-width: 9

Klf4_Int_mppm <- mppm(Klf4 ~ x, random = ~ x | DPI, data = PointPatterns)

summary(Klf4_Int_mppm)
```
The results of this possion regression are present in the log scale. Here, we see the Intercept (8.7), the baseline of KLF4 log-intensity at 0 DPI.The slope (-0.62) indicates that the spatial intensity of KLF4 changes as we move along the x-axis. The negative sign denotes that as x increases, the spatial intensity of KLF4 decreases. Now, the random effects coefficients account for variations that occur at different injury stages. We see that 3D is -0.58, indicating a lower baseline intensity for KLF4 at this stage, while at 14D we see a higher baseline intensity (0.57). The slopes in the random intercepts indicate the adjuste in the rate change as we move across the x-axis. At 0D,a positive adjustment (e.g., 0.30) denotes that the decrease in KLF4 intensity is less pronounced or even reversed at this stage. On the other hand, the negative adjustment at 14D (-0.89) means the decrease is more pronounced.

Overall, these results show a more pronounced increase of KLF4 at 14D and 7D as we move away from the infarct zone, indicating an aggregation of KLF4 in the ischemic zones.


## Generate table for mean intensity

Here, we extract the mean spatial intensity of each point pattern to perform a regression.

```{r}
#| label: tbl-Klf4_ExtractIntensity
#| include: true
#| warning: false
#| message: false

# Initialize an empty data frame to store the results
Klf4_Intensity <- data.frame(MouseID=character(), DPI=factor(), Intensity=integer(), stringsAsFactors = FALSE)

# Iterate over the rows of the hyperframe to calculate intensity for each Klf4 pattern
for (i in 1:nrow(PointPatterns)) {
  # Calculate the intensity of the Klf4 point pattern for the current row
  current_intensity <- summary(PointPatterns$Klf4[[i]])$intensity
  
  # Round the intensity value to have no decimals
  rounded_intensity <- round(current_intensity)
  
  # Extract MouseID and DPI2 for the current row
  current_mouse_id <- PointPatterns$ID[i]
  current_dpi2 <- PointPatterns$DPI[i]
  
  # Add the results to the data frame
  Klf4_Intensity  <- rbind(intensity_results, data.frame(MouseID=current_mouse_id, DPI2=current_dpi2, Intensity=rounded_intensity))
}

Klf4_Intensity$DPI2 <- factor(Klf4_Intensity$DPI2, levels = c("0D", "3D", "7D", "14D", "30D") )

# View the final data frame
print(Klf4_Intensity )

write.csv(Klf4_Intensity , "Data_Processed/Wide10x_Ipsilateral_Klf4_Intensity.csv", row.names = FALSE)
```

## Exploratory data visualization

We conduct exploratory data visualization to observe the distribution of the variables and have first ideas about modeling strategies.

```{r}
#| label: fig-Klf4_Int_Expl
#| include: true
#| warning: false
#| message: false
#| results: false
#| fig-cap: Exploratory data visualization for Klf4 expression
#| fig-width: 9
#| fig-height: 4

# Load the data set in case is not present in the environment

Klf4_Intensity <- read.csv("Data_Processed/Wide10x_Ipsilateral_Klf4_Intensity.csv", header = TRUE)

set.seed(8807)

# Density plot
################

Klf4_Int_Dens <- 
  ggplot(
    data  = Klf4_Intensity, 
    aes(x = Intensity)
    ) +
  geom_density(size = 1.5) +
  geom_rug(size = 1) +
  scale_x_continuous(name ="KLF4 (spatial intensity)") +
  scale_y_continuous(name = "Density") +
  Plot_theme


# Scatter plot
##################

Klf4_Int_Sctr <- 
  ggplot(
    data  = Klf4_Intensity, 
    aes(x = DPI2, 
        y = Intensity)
    ) +
  geom_boxplot() +

scale_y_continuous(name = "KLF4 (spatial intensity)") +
scale_x_discrete(
  name   ="Days post-ischemia (DPI) "
  ) +
Plot_theme

#Plot the result
Klf4_Int_Dens |  Klf4_Int_Sctr 
```

@fig-Klf4_Int_Expl shows a single peak with a skewed distribution. On the right, we see that KLF4 decreases following injury, and starts to rise from 7 days after the injury. The data suggest the peak of KLF4 expression occurs at 14 DPI, when the glial scar is mature.Based on this preliminary visualization, we perform Bayesian modeling having DPI as a categorical variable.

## Statistical modeling for KLF4 spatial intensity

Given the wide distribution of the data un uneven variances, we fit a model with heteroskedasticity (predicting sigma) using a student-t distribution to reduce the impact of extreme values at both extremes. The model takes the following notation:

$$
\mu = \beta_{DPI} \cdot DPI \\
\log(\sigma) = \gamma_{DPI} \cdot DPI 
$$

Where $\beta_{DPI}$ represents the coefficient for `DPI` and $\gamma_{DPI2}$ indicates the coefficient for `DPI` in the model for sigma. With this formulation, we capture both the central tendency and variability of `Intensity` as functions of `DPI2`.

The model takes the following priors: $$
\begin{align}
\beta_{1} \sim Normal(5000,1000) \\
\sigma \sim Student-t(3, 0, 1000), \sigma > 0 \\
\end{align}
$$ \### Fit the models

We employ `brms` to fit the model.

```{r}
#| label: Klf4_Int_Model
#| include: true
#| warning: false
#| message: false
#| results: false
#| cache: true

# Model 1: DPI as a linear predictor
###########################################
Klf4_Int_Mdl1 <- bf(Intensity ~ 0 + DPI2,
                    sigma ~ 0 + DPI2)

get_prior(Klf4_Int_Mdl1, Klf4_Intensity, family = student)

Klf4_Int_Prior1 <- 
  c(prior(normal(5000,1000), class = b, lb= 0),
    prior(student_t(3, 0, 1000), class = b, dpar = "sigma"))

# Fit model 1
Klf4_Int_Fit1 <- 
  brm(
    family = student,
    data    = Klf4_Intensity,
    formula = Klf4_Int_Mdl1,
    prior   = Klf4_Int_Prior1,
    chains  = 4,
    cores   = 4,
    warmup  = 2500, 
    iter    = 5000, 
    seed    = 8807,
    control = list(adapt_delta = 0.99, max_treedepth = 15),
    file    = "Models/Klf4_Expression/Klf4_SpatialInt_Fit1.rds",
    file_refit = "never")
```

### Model diagnostics

To evaluate sample predictions, we perform the model diagnostics using the `pp_check` (posterior predictive checks) function from `brms`. In the graph, ð˜º shows the data and y\~ the simulated data.

```{r}
#| label: fig-Klf4_Int_Diagnostics
#| include: true
#| warning: false
#| message: false
#| results: false
#| fig-cap: Model diagnostics for KLF4 expression
#| fig-height: 4
#| fig-width: 5

set.seed(8807)

Klf4_Int_Fit1_pp <- 
  brms::pp_check(Klf4_Fit1, 
                 ndraws = 100) +
  geom_density(lwd = 2) +
  labs(title = "Posterior predictive checks",
  subtitle = "Formula: Intensity ~ 0 + DPI2, sigma ~ 0 + DPI2") +
  scale_x_continuous(limits=c(0, 20000)) +
  Plot_theme  
  
Klf4_Int_Fit1_pp
```

We see that the model predictions do not deviate meaningfully from the observations.

### Model results

After validating the model, we plot the full posterior distributions:

```{r}
#| label: fig-Klf4_Int_Posterior
#| include: true
#| warning: false
#| message: false
#| results: false
#| fig-cap: Posterior distribution for Klf4 expression
#| fig-width: 9
#| fig-height: 4

set.seed(8807)

Klf4_Intensity_Fig <- Klf4_Intensity %>%
  data_grid(DPI2) %>%
  add_predicted_draws(Klf4_Int_Fit1) %>%
  ggplot(aes(x = DPI2, y = .prediction)) +
  stat_halfeye(alpha = .7) +
  geom_jitter(data = Klf4_Intensity, 
             aes(x = DPI2, y = Intensity),
             width = 0.2,
             size = 1)+
  scale_x_discrete(name = "DPI") +
  scale_y_continuous(name = "KLF4 (spatial intensity)",
                     limits = c(0, 12000),
                      breaks = seq(0, 12000, 3000) ) +
  theme_classic() +
  coord_flip() +
  
  Plot_theme

ggsave(
  plot     = Klf4_Intensity_Fig, 
  filename = "Plots/Klf4_Intensity_Fig.png", 
  width    = 11, 
  height   = 9, 
  units    = "cm")

Klf4_Intensity_Fig 
```

@fig-Klf4_Int_Posterior shows the full posterior distribution for the spatial intensity of KLF4 across time points. We can observe the distribution is specially wide at 14 DPI, where the average value reaches the highest point.

On the other hand, we can also plot the posterior distribution for the sigma parameter:

```{r}
#| label: fig-Klf4_Int_PosteriorSigma
#| include: true
#| warning: false
#| message: false
#| results: false
#| fig-cap: Posterior distribution for Klf4 expression (sigma)
#| fig-width: 9
#| fig-height: 4

# We convert the estimates to a data frame
Klf4_Int_Sigma <- 
  conditional_effects(Klf4_Int_Fit1, dpar = "sigma")

Klf4_Int_Sigma_Fig <- 
  plot(Klf4_Int_Sigma)[[1]] +
  scale_x_discrete(name = "DPI") +
  scale_y_continuous(name = "KLF4 spatial intensity (sigma)",
                     limits = c(0, 4000),
                     breaks = seq(0, 4000, 1500)) +
  theme_classic() +
  coord_flip() +
  Plot_theme

ggsave(
  plot     = Klf4_Int_Sigma_Fig, 
  filename = "Plots/Klf4_Int_Sigma_Fig.png", 
  width    = 10, 
  height   = 8, 
  units    = "cm")


Klf4_Int_Sigma_Fig 
```

We observe that we experience a wide uncertainty at 14 DPI when compared to other time points. We speculate that this uncertainty can be explained by brain shrinkage. Later, we will model that assumption. For the moment, we consider more informative to plot relevant contrast to see the dynamics of KLF4 regulation across time points. To achieve this objective, we use the `emmeans` package [@emmeans].


```{r}
#| label: Klf4_Int_Contrast
#| include: true
#| warning: false
#| message: false
#| results: false

# We generate a data frame with the contrast
Klf4_Int_Contrast <- Klf4_Int_Fit1%>%
emmeans(~ DPI2, var = "Intensity", epred = TRUE) %>%
contrast(method = "revpairwise") %>%
gather_emmeans_draws() %>% sample_n(100)

# Subset relevant data fo the contrast graph
Klf4_Int_Contrast_Sub <- Klf4_Int_Contrast[
  (Klf4_Int_Contrast$contrast=="3D - 0D"|
   Klf4_Int_Contrast$contrast=="7D - 3D"|
   Klf4_Int_Contrast$contrast=="14D - 7D"|
   Klf4_Int_Contrast$contrast=="30D - 14D"|
  Klf4_Int_Contrast$contrast=="30D - 0D"),]

Klf4_Int_Contrast_Sub$contrast <- 
  factor(Klf4_Int_Contrast_Sub$contrast, 
         levels = c("3D - 0D", "7D - 3D", "14D - 7D", "30D - 14D", "30D - 0D")) 
```

Next, we generate the graph:

```{r}
#| label: Klf4_Int_ContrastGraph
#| include: true
#| warning: false
#| message: false
#| results: false

Klf4_Int_Contrast_Fig <- 
  Klf4_Int_Contrast_Sub %>%
  ggplot(
    aes(x    = .value, 
        y    = contrast,
        fill = after_stat(abs(x) < 800)
        )) +
  
  stat_slab() +
  
  geom_vline(xintercept = c(-800, 800), linetype = "dashed") +
  
  stat_pointinterval(
    point_interval = mode_hdi, 
    position = position_dodge(width = .95, preserve = "single")) +
  
    scale_fill_manual(
    name="ROPE", 
    values = c("gray80", "skyblue"),
    labels = c("False", "True")) +
  
  scale_y_discrete(
    name= "",
    labels = c("3D - 0D", "7D - 3D", "14D - 7D", "30D - 14D", "30D - 0D")) +
  
  scale_x_continuous(
   name = "KLF4 spatial intensity (contrast)",
   limits=c(-3000, 4000), 
   breaks=seq(-3000,4000, 2000)) +
  Plot_theme +
  
  theme (legend.position = c(0.8, 0.8))

ggsave(
  plot     = Klf4_Int_Contrast_Fig , 
  filename = "Plots/Klf4_Int_Contrast_Fig.png", 
  width    = 12, 
  height   = 9, 
  units    = "cm")

Klf4_Int_Contrast_Fig

```
### Posterior summary

Next, we plot the posterior summary using the `describe_posterior` function from `bayestestR` package [@bayestestR; @makowski2019].

```{r}
#| label: Klf4_Int_DescribePosterior
#| include: true
#| warning: false
#| message: false
#| results: false
#| cache: true

describe_posterior(
  Klf4_Int_Fit1,
  effects = "all",
  test = c("p_direction", "rope"),
  component = "all",
  centrality = "median")
```
The table displays the regression coefficients with out a base (intercept value). Please note that coefficients for sigma are on in the log scale.


# Analysis of KLF4+ cell intensity


PONER TODOS EN UNA LISTA PARA PLOT MÃS FÃCIL. TENGO QUE CORRER LAS REGRESIONES OTRA VEZ CON ESA NUEVA TABLA

We scale the coordinates

```{r}

library(dplyr)

Klf4_Data <- Klf4_Data %>%
  group_by(MouseID) %>%
  mutate(Scaled_CenterX = (CenterX - min(CenterX)) / (max(CenterX) - min(CenterX)),
         Scaled_CenterY = (CenterY - min(CenterY)) / (max(CenterY) - min(CenterY))) %>%
  ungroup()

Klf4_Data$MouseID <- factor(Klf4_Data$MouseID)
```

We load the `Data_Raw/Widefield_10x_Klf4_Detections/Widefield_10x_Ipsilateral_Klf4_Cells.csv` dataset to very its content.

```{r}
#| label: tbl-Klf4_Cells
#| include: true
#| warning: false
#| message: false
#| tbl-cap: "Data set"

# We load the dataset in case is not present in the R environment
Klf4_Raw <- read.csv(file = "Data_Raw/Widefield_10x_Klf4_Detections/Widefield_10x_Ipsilateral_Klf4_Image.csv", header = TRUE)

knitr::kable(Klf4_Raw [1:10, ])
```

FOR SAMPLING 10% OF THE CELLS

```{r}
# Sample 10% of points per MouseID
Klf4_Data <- Klf4_Data %>%
  group_by(MouseID) %>%
  # Sample 10% within each group
  sample_frac(0.1) %>%
  # Ungroup to remove the grouping structure
  ungroup()
```

## Exploratory data visualization

Next, we visualize the raw data to guide the statistical modeling. We plot the response variable as a density and a scatter plot (per DPI). In the scatter plot, we fit lines for a lineal (black), 2-degree (red), and 3-degree (green) polynomial models.

```{r}
#| label: fig-Klf4_Int_Expl
#| include: true
#| warning: false
#| message: false
#| results: false
#| fig-cap: Exploratory data visualization for Klf4 expression
#| fig-width: 9
#| fig-height: 4

set.seed(8807)

# Density plot
################

Klf4_Int_Dens <- 
  ggplot(
    data  = Klf4_Data, 
    aes(x = Intensity)
    ) +
  geom_density(size = 1.5) +
  geom_rug(size = 1) +
  scale_x_continuous(name ="KLF4 (Intensity)") +
  scale_y_continuous(name = "Density") +
  Plot_theme


# Box plot
##################

Klf4_Int_Box <- 
  ggplot(
    data  = Klf4_Data, 
    aes(x = DPI, 
        y = Intensity)
    ) +
  geom_boxplot()+

Plot_theme

#Plot the result
Klf4_Int_Dens |  Klf4_Int_Box 
```

@fig-Klf4_Int_Expl shows a range of \~ 0 - 10 in the KFL4 intensity. Without accounting for days-post ischemia (DPI), the response variable exhibits a single concentrated peak. On the other hand, the scatter shows that there are no substantial differences between the fitted lines. We speculate that the median (per object) intensity of KLF4 does not vary meaningfully by DPI.

# Statistical modeling for Klf4 expression

We use the `brms` package [@brms; @brms-2] to run Bayesian modeling. We employ weakly informative priors in all the cases to optimize the Markov chain Monte Carlo (MCMC) and the exploration of the parameter space. Given the exploratory data visualization, we fit a linear model to explore the variability of KLF4 by DPI. We build a multilevel model to take into consideration the clustering of cells per animal:

-   **Klf4_Int_Mdl1:** We use `DPI` as a linear predictor for PDGFR-Î² integrated density. The model takes the following notation:

$$
Intensity_{ij} = \beta_{0} + \beta_{1} DPI_{i} + b_{0j} + \epsilon_{i}
$$ Where,

$Intensity_{ij}$ is the intensity measurement for the i-th observation in mouse j. $\beta_{0}$ is the global intercept, $\beta_{1}$ is the fixed effect coefficient of DPI, denoting the average change in intensity for one unit increase in DPI, $b_{0j}$ is the random intercept for mouse, computing the deviation of mouse average intensity from the global average, and $\epsilon_{i}$ is the residual error term.

This model uses the default flat prior from `brms`:

$$
\begin{align}
\beta_{1} \sim Normal(0, 1) \\
\alpha \sim Student-t(3, 500, 300) \\
\sigma \sim Student-t(3, 0, 100), \sigma > 0 \\
\end{align}
$$

## Fit the models for Klf4 intensity

```{r}
#| label: Klf4_Int_Model
#| include: true
#| warning: false
#| message: false
#| results: false
#| cache: true

# Model 1: DPI as a linear predictor
###########################################
Klf4_CellInt_Mdl1 <- bf(Intensity ~ 0 + DPI + (1 | MouseID))

get_prior(Klf4_CellInt_Mdl1, Klf4_Data, family = hurdle_lognormal())

Klf4_CellInt_Prior1 <- 
  c(prior(normal(0,1), class = b, lb = 0),
    prior(student_t(3, 0, 2), class = sd, lb= 0),
    prior(student_t(3, 0, 2), class = sigma, lb=0))

# Fit model 1
Klf4_CellInt_Fit1 <- 
  brm(
    family = hurdle_lognormal(),
    data    = Klf4_Data,
    formula = Klf4_CellInt_Mdl1,
    prior   = Klf4_CellInt_Prior1,
    chains  = 4,
    cores   = 4,
    warmup  = 2500, 
    iter    = 5000, 
    seed    = 8807,
    control = list(adapt_delta = 0.99, max_treedepth = 15),
    file    = "Models/Klf4_Expression/Klf4_CellInt_Fit1.rds",
    file_refit = "never")

                     
#Klf4_CellInt_Fit1 <- 
#  add_criterion(Klf4_CellInt_Fit1, c("loo", "waic", "bayes_R2"))


# Model 2: DPI and x coordinates as linear predictors
###########################################

Klf4_CellInt_Mdl2 <- bf(Intensity ~ 0 + DPI * Scaled_CenterX + (1 | MouseID))

get_prior(Klf4_CellInt_Mdl2, Klf4_Data, family = hurdle_lognormal())

Klf4_CellInt_Prior2 <- 
  c(prior(normal(0,1), class = b),
    prior(student_t(3, 0, 2), class = sd, lb= 0),
    prior(student_t(3, 0, 2), class = sigma, lb=0))

# Fit model 1
Klf4_CellInt_Fit2 <- 
  brm(
    family = hurdle_lognormal(),
    data    = Klf4_Data,
    formula = Klf4_CellInt_Mdl2,
    prior   = Klf4_CellInt_Prior2,
    chains  = 4,
    cores   = 4,
    warmup  = 2500, 
    iter    = 5000, 
    seed    = 8807,
    control = list(adapt_delta = 0.99, max_treedepth = 15),
    file    = "Models/Klf4_Expression/Klf4_CellInt_Fit2.rds",
    file_refit = "never")

#Klf4_CellInt_Fit2 <- 
#  add_criterion(Klf4_CellInt_Fit2, c("loo", "waic", "bayes_R2"))
```

# References

::: {#refs}
:::
